\section*{ПРИЛОЖЕНИЕ Б}
\addcontentsline{toc}{section}{ПРИЛОЖЕНИЕ Б}

\begin{Verbatim}[fontsize=\footnotesize]
Hi. Thanks for your interest in Box! I have answered your question below,
to make things easier to read.
As a general design principle for Box, I try to get the Dynarec not too
complex, but still trying to have decent translated code. Also, I try to be
able to get back to an x86 instruction from an ARM generated code to be
able to handle Signal properly.

Sebastien.

Le mer. 8 d=C3=A9c. 2021 =C3=A0 13:34, Mikhail Nitenko <mnitenko@gmail.com>=
a =C3=A9crit :

> Hello!
>
> I=E2=80=99m a final-year student interested in dynamic binary translation=
,
> specifically in x86 to ARM. I am writing a bachelor thesis about it
> and wanted to explore the current optimizations related to
> translation and wanted to write to you since your project seems to be
> quite advanced.
>
> I read the =C2=ABInner workings=C2=BB and =C2=ABA deep dive into library =
wrapping=C2=BB
> articles and wondered about some things, specifically:
> * How do you determine where the translation block ends? You said
> =C2=ABthere are a lot of possibilities here=C2=BB, from reading QEMU docs=
I
> would imagine that one of them would be a CPU state change that is
> impossible to determine ahead of time.
>
A block is a contiguous array of x86 instructions. A block ends when the
last instruction has no "next" instruction (like a jump, call or ret) , and
when that instruction is not referenced by a jump from the current block.
There are a few exception: multi-bytes NOP are handled, and an unknown
instruction stop the block unconditionally

* How much memory is allocated for each x86 instruction during pass 0?
> Is there some kind of table that determines that?
>
The memory is dynamically allocated in pass 0. So it will use as much
memory as needed.

* JITs (mono) and self-modifying code. I know that emulating
> self-modifying code in x86 is difficult, QEMU for example detects
> writes to pages and invalidates translated code at that page and all
> pages that follow, are you doing the same? Is there some document that
> I can read to learn about problems related to JIT (mono, which I
> assume is open-source .NET)?
>
I'm doing something similar, yes. All pages from which x86 code has been
translated are write protected. Once a write occurs, all blocks generated
from this page are marked dirty, and the jump table from those blocks are
invalidated (to avoid automatic jump from block to block to this one). Once
x86 code wants to run a "dirty" block, a crc32 of the memory is runned and
compared to the crc computed when creating the block. depending if it's the
same or non, the block is marked as clean (and the page protected again) or
deleted and a new one is generated (and the page also protected again)

* Am I correct in understanding that dynarec is used to translate x86
> instructions into ARM instructions and x86 emulation is code in C that
> emulates an instruction?
>
Box86 and Box64 include an Interpreter, coded in C, that can interpret x86
code on any architecture, and a Dynarec, actually only available for ARM.
The Dynarec is coded in C also, with very little assembly file (just the
prolog / epilog of a function call basically). The Dynarec will use
"Emitter", (ab)using C macro, to generate actual arm opcodes.

>
>
> Also, maybe you would be able to point me to some other resource
> where I learn more about translation optimizations? Any help would be
> greatly appreciated!
>
You should have a look at FEX project. This project has a huge emphasis on
the dynarec and code translation optimisation
\end{Verbatim}

\pagebreak
